CODE CHECK

### 2020-06-11 codecheck-paper.Rmd
Update codecheck-paper.Rmd

- **Principles**: https://codecheck.org.uk/
- relation to "proper" citation of data and software, and depositing of data and software in suitable repos _besides_ the bundle for CODE CHECK: do it, only provide the concrete analysis script for the check
- **Why is it useful?**
  - reduce the barrier to evaluating non-text parts of research
  - acknowledges different skill sets and existing high load on reviewers
  - CC breaks problem that publishers staff expertise today is not likely to suffice
  - engaging ECRs is a perfet fit (they are up-to-date in methods, are introduced to peer review)
  - relation to research compendia? point it out for authors unsure about how to structure their workflow
  - CrossRef and Publons and ORCID have features around reviews
  - shifting burden from reviewer to author (but not too much)
- keynote https://doi.org/10.7557/5.4963: in discussion, the speaker mentioned value of peer review for "baseline reassurance", i.e. a paper has at least been checked by someone with an understanding, to increase trust especially if looking at papers from other disciplines
- **report** must fulfil some requirements to make clear what is really checked and not mislead - readers will have very different understandings of a green checkmark!
- reproducibility is hard https://twitter.com/annakrystalli/status/1144176149859377152?ref_src=twsrc%5Etfw > code check gives at least some benefits
- what are the requirements / level or reproducibility required or checked?
  - [Stodden et al., 2013](https://stodden.net/icerm_report.pdf): five-level hierarchy of research - reviewable, replicable, confirmable, auditable, and open or reproducible
  - [Whitaker 2017](https://figshare.com/articles/Showing_your_working_A_guide_to_reproducible_neuroimaging_analyses/4244996/2): reproducible, generalisable, replicability, robustness

@@

- CODECHECK vs. the "Artifacts Evaluated - Functional" badge of the ACM, see https://www.acm.org/publications/policies/artifact-review-badging
- https://paperpile.com/shared/rVNwBS
- Are "Replicated Computations Results (RCR) Report" a thing? apparently so (using ACM terminology)
  - https://dl.acm.org/doi/10.1145/3185337
  - http://users.iems.northwestern.edu/~nelsonb/Publications/Nelson_RCRgreenSim.pdf
  - https://dl.acm.org/doi/epdf/10.1145/3341094
    - citation for replicated workflow lists it as "to appear"
  - getting proper papers for reproductions != CODECHECK approach where we want credit as reviewers
- https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07 - work by Amy J. Ko: https://faculty.washington.edu/ajko/publications
  - is CODE CHECK's process too small, too incremental a change?
- overview of peer review and bibliometrics (if not a reference itself, it gives plenty of references): http://juser.fz-juelich.de/record/865096/files/Peer%20Review%20and%20Bibliometrics.pdf
- _Successes and struggles with computational reproducibility: Lessons from the Fragile Families Challenge_ > https://osf.io/preprints/socarxiv/g3pdb/
- community standards for software, cf. slide 6 in https://figshare.com/articles/_/5675104
  - "good enough practice" is "get a colleague to try using it", slide 19 in https://figshare.com/articles/_/5675104
- journals doing code reviews
  - _"We know of 16 journals that take this valuable step (ie TOP Level 3): AEA as already mentioned, plus others in polisci, psychology, biostats, and org chem: see Data Rep Policies here https://t.co/plOF8j6ADU "_ (https://twitter.com/EvoMellor/status/1202692360456589339?s=09); see also whole thread!
  - https://medium.com/@NeurIPSConf/call-for-papers-689294418f43
  - https://www.journals.elsevier.com/information-systems/editorial-board/
  - SIGMOD Reproducibility Review?
  - Biostatistics "AER" (https://academic.oup.com/biostatistics/article/10/3/405/293660)
- open science platforms/cloud services > https://www.nature.com/articles/d41586-019-03366-x
- live code in articles > o2r, eLife RDS https://www.nature.com/articles/d41586-019-00724-7, ...
- Artefact Evaluation - https://www.artifact-eval.org/about.html
- continuous analysis - https://www.nature.com/articles/nbt.3780
- "A Universal Identifier for Computational Results" > https://www.sciencedirect.com/science/article/pii/S1877050911001256
- Discussion about using Docker containers during JOSS reviews > https://github.com/openjournals/joss/issues/498#issuecomment-462046912
- https://scigen.report, an independent site for registering reproductions, via https://annakrystalli.me/talks/ro-reprohack.html#32
  - example https://scigen.report/reviews/get?doi=10.1063/1.1823034
  - seems to be closed infrastructure
- new journal _ReScienceX_ (about experiments) also seems to have an ECR angle, https://www.nature.com/articles/d41586-020-01328-2
- really interesting work on reproductions in hydrology, with a neat survey that reproducers filled out which might be interesting to capture from codecheckers, too: https://www.nature.com/articles/sdata201930#Sec9

@@

- need publishers build up data science experts who can conduct checks?
- integration of software/data citation checks benefitial?
- bot-support and tools (linter)
- how can open platforms be used to enable a review in private (cf. https://www.cambridge.org/core/blog/2019/08/19/how-to-make-the-data-and-code-for-your-manuscript-available-to-peer-reviewers-before-making-it-public/), if that is desired
- Does nudging towards better practices work (highlight the successes) or is any check with opt-in doomed to fail?
- Do we need more than one codechecker? Is there a similar diversity as in a peer review where more opinions are needed?
- can CODE CHECK work for preprints? how? is a re-check required?
- risk: is the hurdle too low? are "levels" of checks are solution to increase transparency of check's level of detail
- code check in the context of innovations and disruptions in scientific publishing
  - https://doi.org/10.1042/ETLS20180172
  - https://doi.org/10.3390/publications7020034
- should every code be checkd, or should the labour be capped by only checking "important" works or at important stages; if yes, isn't peer review such a stage?
- see https://github.com/codecheckers/discussion
- can we establish code checks as a type of review within the schol comm metadata (or should we?) > we should not, put them on the same level as reviews!
- how can we turn a CODECHECK into a positive experience?
  - cf. kindness in peer review in https://www.mdpi.com/2304-6775/8/2/26

>>>>

- **Principles**: https://codecheck.org.uk/
- relation to "proper" citation of data and software, and depositing of data and software in suitable repos _besides_ the bundle for CODE CHECK: do it, only provide the concrete analysis script for the check
- **Why is it useful?**
  - reduce the barrier to evaluating non-text parts of research
  - acknowledges different skill sets and existing high load on reviewers
  - CC breaks problem that publishers staff expertise today is not likely to suffice
  - engaging ECRs is a perfet fit (they are up-to-date in methods, are introduced to peer review)
  - relation to research compendia? point it out for authors unsure about how to structure their workflow
  - CrossRef and Publons and ORCID have features around reviews
  - shifting burden from reviewer to author (but not too much)
  - What issues with (open) peer review does CODECHECK address?
- keynote https://doi.org/10.7557/5.4963: in discussion, the speaker mentioned value of peer review for "baseline reassurance", i.e. a paper has at least been checked by someone with an understanding, to increase trust especially if looking at papers from other disciplines
- **report** must fulfil some requirements to make clear what is really checked and not mislead - readers will have very different understandings of a green checkmark!
- reproducibility is hard https://twitter.com/annakrystalli/status/1144176149859377152?ref_src=twsrc%5Etfw > code check gives at least some benefits
- what are the requirements / level or reproducibility required or checked?
  - [Stodden et al., 2013](https://stodden.net/icerm_report.pdf): five-level hierarchy of research - reviewable, replicable, confirmable, auditable, and open or reproducible
  - [Whitaker 2017](https://figshare.com/articles/Showing_your_working_A_guide_to_reproducible_neuroimaging_analyses/4244996/2): reproducible, generalisable, replicability, robustness

@@

- CODECHECK vs. the "Artifacts Evaluated - Functional" badge of the ACM, see https://www.acm.org/publications/policies/artifact-review-badging
- https://paperpile.com/shared/rVNwBS
- Are "Replicated Computations Results (RCR) Report" a thing? apparently so (using ACM terminology)
  - https://dl.acm.org/doi/10.1145/3185337
  - http://users.iems.northwestern.edu/~nelsonb/Publications/Nelson_RCRgreenSim.pdf
  - https://dl.acm.org/doi/epdf/10.1145/3341094
    - citation for replicated workflow lists it as "to appear"
  - getting proper papers for reproductions != CODECHECK approach where we want credit as reviewers
- https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07 - work by Amy J. Ko: https://faculty.washington.edu/ajko/publications
  - is CODE CHECK's process too small, too incremental a change?
- overview of peer review and bibliometrics (if not a reference itself, it gives plenty of references): http://juser.fz-juelich.de/record/865096/files/Peer%20Review%20and%20Bibliometrics.pdf
- Open Peer Review > https://researchintegrityjournal.biomedcentral.com/articles/10.1186/s41073-019-0063-9
  - Are OPR lessons transferable to CODECHECK? Does CODECHECK work better with OPR than with (single/double) blind peer review?
  - similarities to their advice: open reports, give credit
- _Successes and struggles with computational reproducibility: Lessons from the Fragile Families Challenge_ > https://osf.io/preprints/socarxiv/g3pdb/
- community standards for software, cf. slide 6 in https://figshare.com/articles/_/5675104
  - "good enough practice" is "get a colleague to try using it", slide 19 in https://figshare.com/articles/_/5675104
- journals doing code reviews
  - _"We know of 16 journals that take this valuable step (ie TOP Level 3): AEA as already mentioned, plus others in polisci, psychology, biostats, and org chem: see Data Rep Policies here https://t.co/plOF8j6ADU "_ (https://twitter.com/EvoMellor/status/1202692360456589339?s=09); see also whole thread!
  - https://medium.com/@NeurIPSConf/call-for-papers-689294418f43
  - https://www.journals.elsevier.com/information-systems/editorial-board/
  - SIGMOD Reproducibility Review?
  - Biostatistics "AER" (https://academic.oup.com/biostatistics/article/10/3/405/293660)
- open science platforms/cloud services > https://www.nature.com/articles/d41586-019-03366-x
- live code in articles > o2r, eLife RDS https://www.nature.com/articles/d41586-019-00724-7, ...
- Artefact Evaluation - https://www.artifact-eval.org/about.html
- continuous analysis - https://www.nature.com/articles/nbt.3780
- "A Universal Identifier for Computational Results" > https://www.sciencedirect.com/science/article/pii/S1877050911001256
- Discussion about using Docker containers during JOSS reviews > https://github.com/openjournals/joss/issues/498#issuecomment-462046912
- https://scigen.report, an independent site for registering reproductions, via https://annakrystalli.me/talks/ro-reprohack.html#32
  - example https://scigen.report/reviews/get?doi=10.1063/1.1823034
  - seems to be closed infrastructure
- new journal _ReScienceX_ (about experiments) also seems to have an ECR angle, https://www.nature.com/articles/d41586-020-01328-2
- really interesting work on reproductions in hydrology, with a neat survey that reproducers filled out which might be interesting to capture from codecheckers, too: https://www.nature.com/articles/sdata201930#Sec9

@@

- need publishers build up data science experts who can conduct checks?
- integration of software/data citation checks benefitial?
- bot-support and tools (linter)
- how can open platforms be used to enable a review in private (cf. https://www.cambridge.org/core/blog/2019/08/19/how-to-make-the-data-and-code-for-your-manuscript-available-to-peer-reviewers-before-making-it-public/), if that is desired
- Does nudging towards better practices work (highlight the successes) or is any check with opt-in doomed to fail?
- Do we need more than one codechecker? Is there a similar diversity as in a peer review where more opinions are needed?
- can CODE CHECK work for preprints? how? is a re-check required?
- risk: is the hurdle too low? are "levels" of checks are solution to increase transparency of check's level of detail
- code check in the context of innovations and disruptions in scientific publishing
  - https://doi.org/10.1042/ETLS20180172
  - https://doi.org/10.3390/publications7020034
- should every code be checkd, or should the labour be capped by only checking "important" works or at important stages; if yes, isn't peer review such a stage?
- see https://github.com/codecheckers/discussion
- can we establish code checks as a type of review within the schol comm metadata (or should we?) > we should not, put them on the same level as reviews!
- how can we turn a CODECHECK into a positive experience?
  - cf. kindness in peer review in https://www.mdpi.com/2304-6775/8/2/26
- how can we allow reactions to checks? (commenting on reports)
